---
title: "Interview Assignment - Data Collection"
author: "Peter Mburu Kimani"
date: "20/04/2022"
output:
  pdf_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    toc_depth: 6
    number_sections: yes
    theme: lumen
---
## Introduction 
This R workbook shows how to retrieve Twitter's Data using Twitter's API Elevated Access. 


## Loading Necessary Packages

```{r}
#creating a vector of the packages that are required
packages<-c("rtweet","tidyverse","httr")

#creating a function that checks if the packages are available in the library, if not they are loaded from CRAN
package.load.install <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```


## Connecting to Twitter's API
After getting Twitter's developer access, the following API keys are used to download the data 
```{r}
key<-"crUhHKQHyP6zjGz6aS5RCAe0U"  #API KEY 
api.secret<-"DFP5kVTtoMEKKWlJB8RrgVvuoxx7lv8UYo0pmXn8VWUlkTy1X2" #API KEYSECRET
access.token<-"2387182982-002DRJGIGduTsQjlVLqXxWhHl6bE3LckJmX53Dk" #ACCESS TOKEN
token.secret<-"NEUH5HRHGvbm81zjNuA06IvjrbTn3AXGmc9HLNpDPiUXo" #TOKEN SECRET
```

#creating twitter token 
```{r}
## Create Twitter token
token <- create_token(
  app = "CFA_Peter",
  consumer_key = key,
  consumer_secret = api.secret,
  access_token =access.token,
  access_secret = token.secret)
```


##Scrapping Twitter 
Scrapping twitter data
```{r}
## search for 18000 tweets using the rstats hashtag
tweet.scrap<- search_tweets(
  "#RacistEU", include_rts = TRUE,retryonratelimit = TRUE,
  type="mixed"
)


```

### Exploring the scrapped data 
Checking how many tweets were downloaded

```{r}
dim(tweet.scrap)
```
26 tweets were downloanded. 

Checking the period range of the downloaded tweet 
```{r}
range(tweet.scrap$created_at)
```
Twitters API Elevated Access, does not give access to tweets that are more than seven days old. Twitter has more restriction to archived tweets (tweets more than 7 days old as the access is given only to Academic Research Track). To overcome this challenge, the tweets for the period 26 February 2022 to 26th March 2022, were obtained from a third party app called Apify (using the free access option) https://console.apify.com/. 

## Exporting scrapped data as CSV
Exporting the scrapped data using API. 
```{r}
#write_as_csv(tweet.scrap,"scrapped_data.csv")
```




